Simple repression draft

### 2020-05-26 doc/section_05_discussion.tex
fix definition of fano factor

Our key insights can be summarized as follows. First, as shown in 
Figure~\ref{fig1:means_cartoons}, the mean expression in the simple repression
architecture is captured by a master curve in which the action of repressor and
the details of the polymerase interaction with the promoter appear separately
and additively in an effective free energy. Interestingly, as has been shown
elsewhere in the context of the Monod-Wyman-Changeux model, these kinds of
coarse-graining results are an exact mathematical result and do not constitute
hopeful approximations or biological naivete \cite{Razo-Mejia2018, Chure2019}.
To further dissect the relative merits of the different models, we must appeal
to higher moments of the gene expression probability distribution. To that end,
our second set of insights focus on gene expression noise, where it is seen that
a treatment of the constitutive promoter already reveals that some models have
Fano factors (standard deviation/mean) that are less than one, at odds with
any and all experimental data that we are aware of.
This theoretical result allows us to directly discard a subset of the models
(models 1-3 in Figure~\ref{fig2:constit_cartoons}) since they cannot be
reconciled with experimental observations. The two remaining models (models 4 and 5
in Figure~\ref{fig2:constit_cartoons}) appear to contain enough microscopic
realistm to be able to reproduce the data. A previous exploration of one 
of these models demonstrated the sloppy nature of the model in which data on single-cell
mRNA counts alone cannot constrain the value of all parameters 
simultaneously \cite{Razo-Mejia2020}. Here we demonstrate that the proposed
one-state bursty promoter model (model 5 in 
Figure~\ref{fig2:constit_cartoons}) emerges as  a limit of the commonly used
two-state promoter model \cite{Peccoud1995}. We put  the idea to the test that this
level of coarse-graining is rich enough to reproduce previous 
experimental observations. In particular we perform Bayesian inference to determine the
two parameters describing the full steady-state mRNA distribution, finding that
the model is able to provide a quantitative description of  a plethora of promoter sequences with different
mean levels of expression and noise.

>>>>

Our key insights can be summarized as follows. First, as shown in 
Figure~\ref{fig1:means_cartoons}, the mean expression in the simple repression
architecture is captured by a master curve in which the action of repressor and
the details of the polymerase interaction with the promoter appear separately
and additively in an effective free energy. Interestingly, as has been shown
elsewhere in the context of the Monod-Wyman-Changeux model, these kinds of
coarse-graining results are an exact mathematical result and do not constitute
hopeful approximations or biological naivete \cite{Razo-Mejia2018, Chure2019}.
To further dissect the relative merits of the different models, we must appeal
to higher moments of the gene expression probability distribution. To that end,
our second set of insights focus on gene expression noise, where it is seen that
a treatment of the constitutive promoter already reveals that some models have
Fano factors (variance/mean) that are less than one, at odds with
any and all experimental data that we are aware of.
This theoretical result allows us to directly discard a subset of the models
(models 1-3 in Figure~\ref{fig2:constit_cartoons}) since they cannot be
reconciled with experimental observations. The two remaining models (models 4 and 5
in Figure~\ref{fig2:constit_cartoons}) appear to contain enough microscopic
realistm to be able to reproduce the data. A previous exploration of one 
of these models demonstrated the sloppy nature of the model in which data on single-cell
mRNA counts alone cannot constrain the value of all parameters 
simultaneously \cite{Razo-Mejia2020}. Here we demonstrate that the proposed
one-state bursty promoter model (model 5 in 
Figure~\ref{fig2:constit_cartoons}) emerges as  a limit of the commonly used
two-state promoter model \cite{Peccoud1995}. We put  the idea to the test that this
level of coarse-graining is rich enough to reproduce previous 
experimental observations. In particular we perform Bayesian inference to determine the
two parameters describing the full steady-state mRNA distribution, finding that
the model is able to provide a quantitative description of  a plethora of promoter sequences with different
mean levels of expression and noise.
